---
title: "STAT 345 Midterm Project"
font: 12pt
date: "Due March 29"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## Golden State Warriors Shot Selections

```{r, echo=FALSE,eval=FALSE}

library(devtools)
devtools::install_github("abresler/nbastatR")


```
    
```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(ggplot2)
library(gridExtra)
library(png)
library(dplyr)
library(magrittr)
library(RColorBrewer)
library(nbastatR)
library(grid)
Sys.setenv("VROOM_CONNECTION_SIZE" = 1000000)
W2004 <- teams_shots(teams = "Golden State Warriors", seasons = 2004, return_message = F)
W2006 <- teams_shots(teams = "Golden State Warriors", seasons = 2006, return_message = F)
W2008 <- teams_shots(teams = "Golden State Warriors", seasons = 2008, return_message = F)
W2010 <- teams_shots(teams = "Golden State Warriors", seasons = 2010, return_message = F)
W2012 <- teams_shots(teams = "Golden State Warriors", seasons = 2012, return_message = F)
W2014 <- teams_shots(teams = "Golden State Warriors", seasons = 2014, return_message = F)
W2016 <- teams_shots(teams = "Golden State Warriors", seasons = 2016, return_message = F)
W2018 <- teams_shots(teams = "Golden State Warriors", seasons = 2018, return_message = F)
W2020 <- teams_shots(teams = "Golden State Warriors", seasons = 2020, return_message = F)
W2022 <- teams_shots(teams = "Golden State Warriors", seasons = 2022, return_message = F)

Wdata <- bind_rows(W2004, W2006, W2008,W2010,W2012,W2014,W2016,W2018,W2020,W2022)


court_img <- "halfcourt.png"
court_img2 <- readPNG(court_img, native = TRUE)
court_plot <- rasterGrob(court_img2, interpolate = TRUE)

base_plot <- ggplot(Wdata, aes(locationX, locationY)) +
  annotation_custom(court_plot, xmin = -290, xmax = 290, ymin = -445, ymax = 70) +
    geom_point(aes(color = isShotMade), alpha = .2,size = .25) +
  xlim(-250, 250) + ylim(500, -10) +
  ggtitle(paste("2003-2022 Golden State Warriors Shot Chart"))+
  xlab("Shot Location")+
  ylab("Shot Location")+
  facet_wrap(vars(slugSeason), nrow = 2)+
  scale_color_manual(values = c("red", "green"), labels = c("Miss", "Make"), name = "Result")

base_plot

```
## Exploration of Shot Chart
   
In these shot charts that I have made for the 2003-2022 seasons I have found some interesting findings.

- The team has faded away from the mid range which makes the shot almost obsolete.
- Shots are becoming more grouped into certain areas of the court such as under the hoop and outside the three point line due to the introduction of efficiency.
- In the years from 2003-2006 the team did shoot many three pointers having most of their shots coming from around the basket and in the mid range.
- In 2007 we start to see more and more three pointers being shot and for the next couple years there were more shots outside of the paint than in previous years. This lets us know that the team feels more comfortable shooting the ball from range.
- The more current years of 2013 to presently there is a consistent pattern of the team shooting deeper three pointers meaning they feel more comfortable from that range.
- In the future it seems that the NBA will keep changing into deeper and deeper to avoid contested shots which in turn makes the playable court bigger and more open. 
   
Over the years basketball has evolved and keeps evolving every year. In this particular case that I am diving into the use of the mid range jump shot and how it is slowly becoming less relevant in comparison to the three pointer. In the graphs above it can be seen how the data points that were once more spread evenly through the court are now being attracted to under the hoop as well as behind the three point line. In the last five to ten years there has been a spike in analytics influencing basketball to a numbers game. For example the reason the mid range jump shot is fading is because players are becoming just as efficient at three pointers which creates a higher points per attempt when shooting more three pointers. The game will keep changing and shot charts are sure to continue to look different over the upcoming years.    

## Code Clean-up
```{r,message=FALSE,warning=FALSE, echo=FALSE}
#import the necessary libraries to run code 
library(ggplot2)
library(gridExtra)
library(png)
library(dplyr)
library(magrittr)
library(RColorBrewer)
library(nbastatR)
library(grid)

#Allow for all the data to be read in by adjusting size capabilites 
Sys.setenv("VROOM_CONNECTION_SIZE" = 1000000)

W2004 <- teams_shots(teams = "Golden State Warriors", seasons = 2004, return_message = F)
W2006 <- teams_shots(teams = "Golden State Warriors", seasons = 2006, return_message = F)
W2008 <- teams_shots(teams = "Golden State Warriors", seasons = 2008, return_message = F)
W2010 <- teams_shots(teams = "Golden State Warriors", seasons = 2010, return_message = F)
W2012 <- teams_shots(teams = "Golden State Warriors", seasons = 2012, return_message = F)
W2014 <- teams_shots(teams = "Golden State Warriors", seasons = 2014, return_message = F)
W2016 <- teams_shots(teams = "Golden State Warriors", seasons = 2016, return_message = F)
W2018 <- teams_shots(teams = "Golden State Warriors", seasons = 2018, return_message = F)
W2020 <- teams_shots(teams = "Golden State Warriors", seasons = 2020, return_message = F)
W2022 <- teams_shots(teams = "Golden State Warriors", seasons = 2022, return_message = F)

#combine all of the years together to make a big data set
Wdata <- bind_rows(W2004, W2006, W2008,W2010,W2012,W2014,W2016,W2018,W2020,W2022)


#import background court image
court_img <- "halfcourt.png"
court_img2 <- readPNG(court_img, native = TRUE)


# Plot the court image
court_plot <- rasterGrob(court_img2, interpolate = TRUE)


# Set up the base plot with court background
# Create the start of the plot with the data and shot locations
base_plot <- ggplot(Wdata, aes(locationX, locationY)) +
  
  #adjust the data and court to fit each other 
  annotation_custom(court_plot, xmin = -290, xmax = 290, ymin = -445, ymax = 70) +

  
  #Adjust the size of the data points 
    geom_point(aes(color = isShotMade), alpha = .2,size = .15) +
  
  
  
  #Adjust where the data points can fall
  xlim(-250, 250) + ylim(500, -10) +
  
  #Add a title and axis labels to the chart 
  ggtitle(paste("2003-2022 Golden State Warriors Shot Chart"))+
  xlab("Shot Location")+
  ylab("Shot Location")+
  
  #change the colors of made and missed shots so it is easily distinguishable 
  facet_wrap(vars(slugSeason), nrow = 2)+
  scale_color_manual(values = c("red", "green"), labels = c("Miss", "Make"), name = "Result")
#Finally show the plot
base_plot
```
   
## Clutch Time Shots

    

```{r,message=FALSE,warning=FALSE,echo=FALSE}
new_data <- filter(Wdata, minutesRemaining <= 2)

#import background court image
court_img <- "halfcourt.png"
court_img2 <- readPNG(court_img, native = TRUE)


# Plot the court image
court_plot <- rasterGrob(court_img2, interpolate = TRUE)


# Set up the base plot with court background
# Create the start of the plot with the data and shot locations
base_plot <- ggplot(new_data, aes(locationX, locationY)) +
  
  
  annotation_custom(court_plot, xmin = -290, xmax = 290, ymin = -445, ymax = 70) +

  
  #Adjust the size of the data points 
    geom_point(aes(color = isShotMade), alpha = .2,size = .15) +
  
  
  
  #Adjust where the data points can fall
  xlim(-250, 250) + ylim(500, -10) +
  
  #Add a title and axis labels to the chart 
  ggtitle(paste("2003-2022 Golden State Warriors Clutch Time Shot Chart"))+
  xlab("Clutch Time Shot Location")+
  ylab("Clutch Time Shot Location")+
  
  #change the colors of made and missed shots so it is easily distinguishable 
  facet_wrap(vars(slugSeason), nrow = 2)+
  scale_color_manual(values = c("red", "green"), labels = c("Miss", "Make"), name = "Result")

base_plot

#Get the clutch time shooting percentage for each year
average_data <- Wdata %>%
  filter(minutesRemaining <= 2) %>%
  group_by(Year = yearSeason) %>%
  summarise(Avg_Clutch_Pct = mean(ifelse(isShotMade == "TRUE", 1, 0)))

# Create a bar chart to visualize the average clutch shooting percentages
bar_chart <- ggplot(average_data, aes(x = Year, y = Avg_Clutch_Pct)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Average Clutch Shooting Percentage by Year", x = "Year", y = "Average Clutch Shooting Percentage") +
  # Set y-axis limits to zoom in on 0.4-0.5 range
  theme_minimal()

# Display the bar chart
print(bar_chart)

```

  In addition to the regular shot chart for the Warriors in the seasons 2004-2022 I chose to include a shot chart and a bar chart for clutch time shots. Clutch time shots are defined as shots taken with two or less minutes in a game. In the clutch time shot chart I explored the shots that our team is taking when we need o make shots to close out games or bring the game back to become winnable. These shots do more more than add points to the board they can swing the momentum and even change a season for teams. In the bar chart we see that The Warriors climbed up in clutch time shooting percentage consistently and saw a big jump in the years 2015-2019. Uncoincidentally the Warriors went to the Finals all of those years and even won three of them. After 2019 they see a drop and then it picks back into the more present years of 2022. This again can be seen in their record and season results as in 2020 they had their worst record in the recent past and then when the percentage jumps up again in 2022 they won another world championship. Possibly most importantly when thinking about the Warriors success in the clutch time is when the shot chart and percentage are paired together we see that more three pointers are taken as well as percentage going up. Combining more three pointers with higher efficiency means more points than ever are being scored in the clutch time for the Warriors helping them win more games and pull more momentum to their side. Their success in the recent past can be attributed to these clutch time stats. 
  
  


